---
title: "estimación de la función de supervivencia para eventos recurrentes"
author: "Andy Vacas Garcia, 1600304"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,eval=T,include=T,message = FALSE,warning=FALSE)
```

```{r}
# Cargar paquetes


library(survival)
library(ggplot2)
devtools::install_github("isglobal-brge/survrec",
build = FALSE)
library(survrec)


install.packages("C:/Users/andyv/Downloads/gcmrec_1.0-5.zip", repos = NULL, type = "win.binary")

devtools::install_github("isglobal-brge/TestSurvRec",build = FALSE)
library(TestSurvRec)
```


Data lymphoma is available at gcmrec package. It contains cancer relapses times after first treatment in patients diagnosed with low grade lymphoma. Data can be loaded into R by executing:

```{r}
data(lymphoma, package = "gcmrec")
```

NOTE: variable time contains inter-occurrence times, event is the censoring variable that is 1 for cancer relapses and 0 for the last follow up time indicating that the event is not observed and the variable id identifies each patient.



Exercise 1:
Estimate survival function using Peña-Strawderman-Hollander, Wang-Chang and a Frailty model.
```{r}
# Peñá-Strawderman-Hollander

fit_psh <- survfitr(Survr(id, time, event)~1,data=lymphoma, type="pena-strawderman-hollander")

# Wang-Chang
fit_wc <- survfitr(Survr(id, time, event)~1,data=lymphoma, type="wang-chang")

# Frailty model (modelo de supervivencia con efecto aleatorio por paciente)
fit_frailty <-survfitr(Survr(id, time, event)~1,data=lymphoma, type="MLEfrailty")
```


Represent the three estimated survival curves in a figure.

```{r}
plot(fit_wc,conf.int = FALSE)
lines(fit_psh,col="red")
lines(fit_frailty,col="green")
legend("topright",c("WC","PSH","Frailty"),col=c("black","red","green"),lty = 1)

```

Is there correlation among inter-occurrence times?

```{r}
# Correlación de los tiempos entre eventos por paciente
library(dplyr)

# Tiempos por paciente
pairs <- lymphoma %>%
  group_by(id) %>%
  filter(n() > 1) %>%
  summarise(cor = ifelse(n() >= 2, cor(time[-n()], time[-1]), NA))

mean(pairs$cor, na.rm = TRUE)  # media de correlación entre eventos
```


Which is the best method to analyze these data?


Exercise 2:
Investigate how the package TestSurvRec compares two whole survival curves (see References section to both package and manuscript describing how it works).

By using this method:

Compare cancer relapse times between males and females (variable sex).

```{r}
subset_00<-lymphoma
subset_00$group<-as.factor(subset_00$sex)
Plot.Event.Rec(subset_00)
# Comparar curvas de supervivencia: sexo
Dif.Surv.Rec(subset_00,"all",1,1,0,0)

Dif.Surv.Rec(subset_00,"Grec")
Print.Summary(subset_00)
```

Compare cancer relapse time between patients having single lesions and localized lesions (variable distrib).
```{r}
subset_01 <- lymphoma[lymphoma$distrib %in% c(0, 1), ]
subset_01$group<-as.factor(subset_01$distrib)

Plot.Event.Rec(subset_01)
# Comparar curvas de supervivencia: sexo
Dif.Surv.Rec(subset_01,"all",1,1,0,0)

Dif.Surv.Rec(subset_01,"Grec")
Print.Summary(subset_01)
```

Compare cancer relapse time between patients having single lesions and lesions in more than one nodal site (variable distrib).

```{r}
subset_13 <- lymphoma[lymphoma$distrib %in% c(1, 3), ]
subset_13$group<-as.factor(subset_13$distrib)

Plot.Event.Rec(subset_13)
# Comparar curvas de supervivencia: sexo
Dif.Surv.Rec(subset_13,"all",1,1,0,0)

Dif.Surv.Rec(subset_13,"Grec")
Print.Summary(subset_13)
```


NOTE: variable distrib encodes the lesions involved at diagnosis and has 4 categories (0 = Single, 1 = Localized, 2 = More than one nodal site, 3 = Generalized)